### 2023-01-26 paper-thesis/chapters/40-discussing-statix.tex
Work on "Ergonomics of Spoofax"

	\item A better way to control the errors generated by Statix, such that users of the language get good errors. Statix does support \texttt{expr | error "Error text here"} annotations, but these don't allow constraints that instantiate metavariables on the left hand side whereas this would be needed for good error messages for dependently typed languages, since \verb|expectBetaEq| may infer variables.
\end{enumerate}

@@

- performance
- testing works well
- specifying grammars is still difficult
- debugging (statix tests are not great)

>>>>

	\item A better way to control the errors generated by Statix, such that users of the language get good errors. Statix does support \texttt{expr | error "Error text here"} annotations, but these don't allow constraints that instantiate metavariables on the left hand side whereas this would be needed for good error messages for dependently typed languages, since \verb|expectBetaEq| may infer variables. \todo{This is not true}
\end{enumerate}

@@

Next we will describe our experience with Spoofax as a whole.

\begin{enumerate}
	\item Spoofax is still slow, even though its performance has improved over the past years. On a high-end computer\footnote{Intel i7 6700k @ 4.2GHz, 16 GB Ram @ 2133MHz, Arch Linux @ Jan 2023, Spoofax 3 v0.19.2}, we took the following measurements.
	\begin{itemize}
		\item Full build (non-incremental) takes 55 seconds
		\item Incremental build changing only one Statix file takes 5 seconds
		\item Running all 171 tests takes 18 seconds
	\end{itemize}
	While these numbers are not unworkably high, they are still magnitudes higher than the implementation in Haskell (chapter \ref{ch:comp-haskell}) and other general-purpose languages. Improving these numbers would make the experience better.

	\item The SPT testing DSL works very well, it makes writing integration tests very easy. However, support for unit tests is still lacking. The to create unit tests is using Statix tests, allowing assertions on specific relations. But these tests are not integrated with SPT tests (they don't even reside in the test folder) and there is no way to run all Statix tests, to ensure no regressions happened. Ideally there would be one button that runs all SPT and Statix tests.

	\item Debugging Statix is still quite tedious. When we have a constraint that does not do what we expect it to do (ie. we have a failing Statix test), one either has to stare at its definition until they figure out the problem, or start unfolding the definitions of the constraints into the Statix test, finding out which part of the definition is misbehaving and repeating the unfolding. Having support for a debugger would help tremendously with this.
\end{enumerate}
